---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data.... \newpage

# Data {#sec-data}

## Source

The original paper, "The Unequal Distribution of Opportunity: A National Audit Study of Bureaucratic Discrimination in Primary School Access(2020) by Olsen, Kyhse-Anderson, Moynihan", was replicated from American Journal of Social Science.

@fig-decision_rate1 is one of the results in the original study, and it shows the decision type rate based on the different treatments (Muslim/Danish). In addition to the results depicted in @fig-decision_rate1, we proceeded the analysis by incorporating new variables, namely the characteristics of schools.

```{r}
#| echo: false
#| label: fig-decision_rate1
#| message: false
#| fig-cap: Decision type rate based on each treatment
knitr::include_graphics(file.path(here::here(),"outputs/models/replicated_figure.png"))
```

## Overview of Senders and Respondents' traits

In the original experiment, the designed sender attributes consist of two categories: names (Muslim and Danish names) and the Diligence trait. Also, the respondent schools are classified based on six characteristics: whether the school is located in a major city, its geographical region, the average student grades, the number of students, the number of non-Danish students, and the school type. Theses school characteristics were investigated through the Danish Ministry of Education database in the original study. Below is a summary.

[***Senders trait***]{.underline}

-   **Name**:

    -   Danish: Peter Nielsen

    -   Muslim: Mohammad Osman

-   **Diligence trait**:

    Diligence trait was classified based on the content of the emails, according to the experiment setting in the original study.

    -   Hi \[school name\]. I am writing because we would like to move our son to another school.

        [Non-diligent student treatment]{.underline}: He is in 3rd grade and we don't think that it is going very well at this school

        [Diligent student treatment]{.underline}: He is in 3rd grade and the teachers say that he is doing well with his homework. but we still don't think that it is going very well at this school

[***Respondents'(School) trait***]{.underline}

-   **Major(Big) city**: whether the school is located in a big city

-   **Geographical region**: Reion Hovedstade, Region Midtjylland, Region Nordjylland, Region Sjaelland or Region Syddanmark

-   **School type**: Public or Private

-   **Number of Students**

-   **Number of Non-Danish Students**

-   **Average Grade of Students**

## Methodology

As mentioned above, this paper examines how the characteristics of schools, in addition to the original experiment's results for decision type rates, impact Muslim's admission to private schools in Denmark. The following subsection will provide detailed explanations of how and why each piece of data was used or not.

### Decision type

As shown on @fig-decision_rate1, in the original study, the schools' responses were categorized into three decision types: clear rejection, unclear response and clear acceptance. According to the study, two coders categorized each response email into these three decision types, and the responses categorized by two coders were denoted by response1 and response2. As a result, the Cronbach's alpha analysis revealed minimal difference in perception between the two coders regarding the response emails. Therefore, in this paper, only Response type 1 is used, and the they are denoted as reject1, unclear1, accept1 in the data set.

### Using Senders and Respondents' traits

Followed by the Senders and Respondents' trait above, the original study also investigated whether each school is above the median level of trait(Number of Students, Number of Non-Danish Students and Students' grade). Based on this examination, this study examined how schools with above-median and below-median traits influence the decision type rate through proportional contrasts. Additionally, the impact of school location and school type on Muslim's decision type rate will be investigated.

### Not using Diligence trait

Finally, diligence trait is excluded from this study. This decision is based on the original research, which revealed through the full ordinary least square(OLS) results that diligence traits do not have any effect on outcome variables such as response rate or the decision to accept or reject students.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References
