---
title: "Bureaucratic discrimination stemming from Islamophobia needs to be highlighted"
author: 
  - Fares Alkorani
  - Kyungrok Park 
  - Sachin Chihikara
thanks: "Code and data are available at: https://github.com/SachinChhikara/administrative_discrimination_analysis.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
```

# Introduction

From the past to the present, bureaucracy has been a characteristic predominantly found in large companies or public institutions. With the nature of bureaucracy, specific individuals or groups within an organization tend to hold the authority to make most administrative decisions, which can result in administrative discrimination stemming from subjective or group opinions or beliefs influencing administrative decisions.

The original study, “The Unequal Distribution of Opportunity: A National Audit Study of Bureaucratic Discrimination in Primary School Acces (2020) by Olsen, Kyhse-Anderson, Moynihan”, experimented with the unequal opportunities brought about by bureaucratic discrimination in Danish primary schools. Before the experiment, Olsen, Kyhse-Anderson, Moynihan defined bureaucratic discrimination as two forms: either through withholding opportunities in service or imposing administrative burdens to make bureaucratic procedures more complex and difficult. As a result, the original study revealed that individuals with Muslim names experience both forms of bureaucratic discrimination more frequently than those with native (Danish) names.

Our paper will reproduce these findings and will focus on the potential administrative discrimination against Muslims in Canada by building on the two forms of administrative discrimination (1.Withholding opportunities in service, 2.Imposing administrative burdens to make bureaucratic procedures more complex and difficult).
Thus, the estimand for our research will be the effect of being Muslims to get admitted to the private school in Denmark (Broadly effect of being Muslims to access the public service) . 
Through the original experiment’s results, we discovered that administrative discrimination against Muslims could be driven by Islamophobic attitudes. Based on the two forms of bureaucratic discrimination defined in the original experiment, our study aims to assess perceptions about Muslims in Canada and discuss the potential for widespread administrative discrimination against Muslims.
In our study, the statistical programming language R software (R Core Team, 2023) was used for replication purposes. Additional packages, tidyverse (Wickham et al.2019), knitr(Xie 2014), ltm(Rizopoulos 2006), grindExtra(Auguie 2017), estimatr, texreg(Leifeld 2013), scales(Wickham et al. 2023), cowplot(O. Wilke 2024), ggplot2 (Wickham 2016) and here(Müller 2020) were also utilized for our further analysis and reproduction. 

This paper is organized into 4 sections: Introduction, Data, Results and Discussion. The Data section provides a description of the variables and experimental setup used in the original paper, including how administrative discrimination was measured. The Results section cites two graphs from the original paper to reaffirm the existence of administrative discrimination against Muslims in Denmark. Finally, in the Discussion section, it is asserted that such administrative discrimination may occur or is occurring in Canada, leveraging Islamophobic sentiments or perceptions of Muslims.

# Data {#sec-data}

## Source

This paper reproduces two results from "The Unequal Distribution of Opportunity: A National Audit Study of Bureaucratic Discrimination in Primary School Access(2020) by Olsen, Kyhse-Anderson, Moynihan" in the American Journal of Social Science. Our study mainly focuses on the findings whether 1) Muslim gets withdrawn from the public service more frequently 2) Muslims gets the the administrative burdens more often than Danish.

## Methodology

The original experiment was conducted with 1,698 primary schools in Denmark, where fictitious emails regarding student transfer requests were sent to each school. These emails were sent under different names (Muslim/Danish), with 850 emails sent under one name and 848 under the other. Fictitious students were also assigned diligence traits, which could be discerned from the content of the emails (Overview of Senders’ and Respondents’ traits). Additionally, the original study examined both the characteristics of the schools and those of the senders. The characteristics of the schools (whether the school is located in a major city, its geographical region, the average student grades, the number of students, the number of non-Danish students, and the school type.) were investigated using the Danish Ministry of Education database, while the characteristics of the senders were determined based on the content of the response emails.
Furthermore, email responses to transfer requests received from schools were measured in terms of decision type rate, compliance cost rate, and psychological cost rate for each ethnicity (Danish/Muslim), depending on the content of the email. For the decision type, response emails were classified into three categories: clear reject, unclear response, and clear accept. Also, for compliance cost, it was categorized into four groups based on the format of the email content: simple question, complex question, request a phone call, and request to meet. Since the difference of decision type rates and compliance cost rates between Danish and Muslim represent two forms of administrative discrimination mentioned above, our study will also verify these two sets of results and discuss the potential administrative discrimination against Muslims in Canada.

### Senders and Respondents'(schools) trait

[***Senders trait***]{.underline}

-   **Name**:

    -   Danish: Peter Nielsen

    -   Muslim: Mohammad Osman

-   **Diligence trait**:

    Diligence trait was classified based on the content of the emails, according to the experiment setting in the original study.

    -   Hi \[school name\]. I am writing because we would like to move our son to another school.

        [Non-diligent student treatment]{.underline}: He is in 3rd grade and we don't think that it is going very well at this school

        [Diligent student treatment]{.underline}: He is in 3rd grade and the teachers say that he is doing well with his homework. but we still don't think that it is going very well at this school

[***Respondents'(School) trait***]{.underline}

-   **Major(Big) city**: whether the school is located in a big city

-   **Geographical region**: Reion Hovedstade, Region Midtjylland, Region Nordjylland, Region Sjaelland or Region Syddanmark

-   **School type**: Public or Private

-   **Number of Students**

-   **Number of Non-Danish Students**

-   **Average Grade of Students**

### Measurement for administrative discrimination

-   ***Type of Administrative Discrimination***

    -   [**Withdrawing Opportunity**]{.underline}

        Categorizing receiving emails by 3 categories

        1.  Clear Reject
        2.  Unclear Response
        3.  Clear Acceptance

        [**Giving administrative burdens**]{.underline}

        Categorizing receiving emails by 4 categories

        1.  Simple Question
        2.  Complex Question
        3.  Request a Phone call
        4.  Request to meet

### Effect of Senders and Respondents' trait

Lastly, the original study revealed through the full ordinary least squares (OLS) results that neither the diligence trait nor the characteristics of schools have any influence on the outcome variables (decision type, compliance cost). Consequently, apart from racial characteristics based on two different names, no other traits will be mentioned in this paper.

# Results

## Decision type rate

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-decision_type
#| fig-cap: "Decision type rate based on each treatment"
knitr::include_graphics(file.path(here::here(),"outputs/models/replicated_figure.png")) 
```

@fig-decision_type illustrates the decision type rate of schools based on two ethnic groups (Danish/Muslim). As evident from the figures, Muslims received approximately 7.4% more clear admission rejections compared to Danes, and unclear responses were also approximately 2.4% higher for Muslims. In contrast, for clear admission acceptance, Danes were approximately 9.8% higher. The higher rate of clear rejections for Muslims compared to Danes demonstrates one form of administrative discrimination, where individuals are initially deprived of opportunities. Conversely, the higher rate of clear admission approvals for Danes compared to Muslims also signifies receiving fewer opportunities in education, a public service, indicating exclusion of Muslims as recipients of public service opportunities due to the first form of administrative discrimination.

## Compliance Cost rate

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-compliance_cost
#| fig-cap: "Compliance cost rate based on each treatment"
knitr::include_graphics(file.path(here::here(),"outputs/models/replicated_figure2.png")) 
```

@fig-compliance_cost displays the compliance cost rate between each ethnic group. Muslims showed higher frequencies in both simple and complex questions, with 8.5% and 0.5% more, respectively. This indicates that Muslims are subjected to discrimination in the form of imposing administrative burdens. Regarding the request for meetings, there was not a significant difference between the two racial groups. However, in contrast to simple and complex questions, Danish individuals showed a 14% higher rate of requesting phone calls. According to the interpretation of the original study results, administrative officers responsible for admissions utilized phone calls as an informal verification method, indicating higher social trust towards Danes. Therefore, the higher rate of phone call requests for Danes does not represent compliance costs but rather reflects higher social trust towards Danes, highlighting administrative discrimination against Muslims.

# Discussion

## Perception about Muslims in Canada and Potential Administrative Discrimination againt them  {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Ethical Reflection
aaa

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References

---
nocite: '@*'
---
